---
title: "p8105_hw2_yl4360"
output: github_document
---

# Problem 1

**Read and clean the Mr. Trash Wheel sheet.**

```{r message = FALSE}
library(tidyverse)
library(dplyr)
library(readxl)
data_MTW = read_excel("./HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N338") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls))
```

**Read and clean precipitation data for 2017 and 2018.**

```{r}
data_2017 = read_excel("./HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2017 Precipitation", range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  drop_na(total)
year_2017 = tibble(year = rep(2017, 13))
data_2017 = bind_cols(year_2017, data_2017)

data_2018 = read_excel("./HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2018 Precipitation", range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  drop_na(total)
year_2018 = tibble(year = rep(2018, 8))
data_2018 = bind_cols(year_2018, data_2018)

data_pre = bind_rows(data_2017, data_2018) %>% 
  mutate(month = month.name[month])

```

**Interpretation**

The `data_MTW` dataset contains `r nrow(data_MTW)` observations of `r ncol(data_MTW)` variables related to dumpster-specific data. The key variables are `r colnames(data_MTW)`.

The `data_pre` dataset contains `r nrow(data_pre)` observations of `r ncol(data_pre)` variables related to each month's total precipitation data for 2017 and 2018. The key variables are `r colnames(data_pre)`. The total precipitation in 2018 is `r data_pre[21, "total"]` and the total precipitation in 2017 is `r data_pre[13, "total"]`.

# Problem 2

**Clean the data in pols-month.csv.**

```{r}
data_pols = read_csv("./fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(col = mon, into = c("year", "month", "day")) %>% 
  mutate(
    month = as.integer(month), 
    month = month.name[month],
    year = as.integer(year),
    day = as.integer(day),
    president = ifelse(prez_gop == 0, "dem", "gop")) %>% 
  select(-prez_gop, -prez_dem, -day)
```

**Clean the data in snp.csv.**

```{r}
data_snp = read_csv("./fivethirtyeight_datasets/snp.csv") %>% 
  separate(col = date, into = c("day", "month", "year")) %>% 
  mutate(
    month = as.integer(month),
    year = as.integer(year),
    day = as.integer(day)) %>% 
  arrange(year, month) %>% 
  mutate(month = month.name[month]) %>% 
  select(year, month, everything())
```

**Tidy the unemployment data.**

```{r}
data_uem = read_csv("./fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(Jan:Dec, 
            names_to = "month",
             values_to = "% of unemployment") %>% 
  rename(year = Year) %>% 
   mutate(month = match(month, month.abb),
         month = month.name[month])

pols_snp = left_join(data_pols, data_snp, by = c("year","month")) 
pols_snp_uem = left_join(pols_snp, data_uem, by = c("year","month"))
```

**Interpretation**

The dataset `data_pols` contains `r nrow(data_pols)` observations of `r ncol(data_pols)` variables related to the number of national politicians who are democratic or republican at any given time. The key variables are `r colnames(data_pols)`. The range of year is from `r data_pols[1, "year"]` to `r data_pols[822, "year"]`.

The dataset `data_snp` contains `r nrow(data_snp)` observations of `r ncol(data_snp)` variables realted to  Standard & Poor’s stock market index (S&P). The key variables are `r colnames(data_snp)`. The range of year is from `r data_snp[1, "year"]` to `r data_snp[787, "year"]`.

The dataset `data_uem` contains `r nrow(data_uem)` observations of `r ncol(data_uem)` variables realted to  the perventage of unemployment at any given time. The key variables are `r colnames(data_uem)`. The range of year is from `r data_uem[1, "year"]` to `r data_uem[816, "year"]`.

# Problem 3

**Load and tidy Popular_Baby_Names.csv.** 

```{r}
baby_names = read_csv("./Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(gender = str_to_lower(gender),
         ethnicity = str_to_lower(ethnicity),
         childs_first_name = str_to_lower(childs_first_name),
         ethnicity = recode(ethnicity, `asian and paci` = "asian and pacific islander", `black non hisp` = "black non hispanic", `white non hisp` = "white non hispanic")) %>% 
  distinct() 
```

**Produce a table showing the rank in popularity of the name “Olivia” as a female baby name over time.**

```{r}
olivia  = baby_names %>% 
  filter(gender == "female", childs_first_name == "olivia") %>% 
  select(year_of_birth, ethnicity, rank) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank") %>% 
  knitr::kable(caption = "Table 1.Tthe rank in popularity of the name “Olivia” as a female baby name over time",align = "c")
olivia
```

**Produce a table showing the most popular name among male children over time.**

```{r}
pop_na_male = baby_names %>% 
  filter(gender == "male", rank == 1) %>% 
  select(-gender, -rank, -count) %>% 
   pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name") %>% 
  knitr::kable(caption = "Table 2. The most popular name among male children over time",align = "c")
pop_na_male
```

**Scatterplot**

```{r}
plot = baby_names %>% 
  filter(
    gender == "male",
    ethnicity == "white non hispanic",
    year_of_birth == 2016) %>% 
  ggplot(aes(x = rank, y = count))+
  geom_point(alpha = 0.5) + 
  labs(title = "Fig 1. Rank of popularity of boy's name in 2016") 

plot
```

